\chapter{Definitions and Acronyms}

% This chapter may change if I figure out how to make the glossary environment work!

\section{Definitions}

\begin{description}
\item[ad] A UML Activity Diagram.  For diagrams generated in Poseidon, this label appears in the
diagram's title box, located at the top left corner of the diagram.
\item[Application] The GMAT executable program.
\item[cd] A UML Class Diagram.  For diagrams generated in Poseidon, this label appears in the
diagram's title box, located at the top left corner of the diagram.
\item[Command] One step in the Mission Control Sequence.
\item[Control Sequence] Either a Mission Control Sequence or a Function Control Sequence.
\item[Engine] The ``guts'' of GMAT, consisting of all of the classes, control structures, objects,
and other elements necessary to run a mission.
\item[Factory or Factories] Components used to create pieces that users use when modeling a mission.
\item[Function Control Sequence] The time ordered steps taken in a GMAT function.
\item[GMAT] General Mission Analysis Tool.
\item[Graphical User Interface] The graphical front end for GMAT, built using the wxWidgets
toolkit.  GMAT can also be built as a console application, but most users work from the GUI.
\item[GUI] The Graphical User Interface.
\item[Interface] The connection between GMAT and external systems, like MATLAB.
\item[Interpreter] The connection point between users and the Application.  GMAT uses a
ScriptInterpreter when constructing a mission from a script file, and a GuiInterpreter when
configuring from the GUI.
\item[Mission] All of the elements configured by the user to solve a specific problem.  Every
element of a GMAT Mission is contained in the Model, but the Model may include components that are
not part of a specific Mission.
\item[Mission Control Sequence] The time ordered steps taken to run a mission.
\item[Model] All of the elements configured by a user in the GMAT Application.
\item[Moderator] The central control point in the Engine.
\item[Parameter] A value or other property calculated outside of a GMAT object.  Parameters as used
in this context are all elements derived from the Parameter base class, as described in
Chapter~\ref{chapter:Parameters}.
\item[Property] A data member of a Resource or Command.  Properties are the internal data associated
with the objects used in a GMAT model.
\item[Resource] An element of the GMAT model that represents an object used when running the Mission
Control Sequence.
\item[Sandbox] The portion of GMAT used to run a mission.
\item[Script] A text file that contains all of the instructions required to configure a mission in
GMAT.
\item[sd] A UML Sequence Diagram.  For diagrams generated in Poseidon, this label appears in the
diagram's title box, located at the top left corner of the diagram.
\item[sm] A UML State Diagram.  For diagrams generated in Poseidon, this label appears in the
diagram's title box, located at the top left corner of the diagram.
\end{description}

\section{Acronyms}

\begin{description}
\item[GMAT] General Mission Analysis Tool
\item[GSFC] Goddard Space Flight Center
\end{description}
