% $Id: SDTools.tex,v 1.1 2008/01/31 18:04:16 dconway Exp $
\chapter{GMAT Software Development Tools}
\chapauthor{Darrel J. Conway}{Thinking Systems, Inc.}

GMAT is a cross-platform mission analysis tool under development at Goddard Space Flight Center and
Thinking Systems, Inc.  The tool is being developed using open source principles, with initial
implementations provided that run on 32-bit Windows XP, Linux, and the Macintosh (OS X).  This
appendix describes the build environment used by the development team on each of these platforms.

The GMAT code is written using ANSI-standard C++, with a user interface developed using the
wxWindows toolkit available from http://www.wxwidgets.org.  Any compiler supporting these standards
should work with the GMAT code base.  The purpose of this document is to describe the tools that
were actually used in the development process.

Source code control is maintained using the Concurrent Versions System (CVS 1.11) running on a
server at Goddard.  Issues, bugs, and enhancements are tracked using Bugzilla 2.20 running on a
server at Goddard.

\section{Windows Build Environment}
\begin{itemize}
   \item Compiler: gcc version 3.4.2 (mingw special)
   \item IDE Tool: Eclipse 3.1.1, with CDT 3.0.1 plug-in
   \item wxWindows Version: wxMSW 2.6.2
\end{itemize}

On Windows, GMAT has also been built using the Dev-C++ environment.

\section{Macintosh Build Environment}
\begin{itemize}
   \item Compiler: gcc 4.0.1, XCode v. 2.2
   \item IDE Tool: Eclipse 3.1.2, with CDT 3.0.1 plug-in
   \item wxWindows Version: wxMac 2.6.2
\end{itemize}

\section{Linux Build Environment}

GMAT is regularly built on two different Linux machines at Thinking Systems, one running Mandriva
Linux, and the second running Ubuntu Linux.  Both build environments are listed here.

\paragraph{On Mandriva 2006}
\begin{itemize}
   \item Compiler: gcc version 4.0.1 (4.0.1-5mdk for Mandriva Linux release 2006.0)
   \item IDE Tool: Eclipse 3.1.1, with CDT 3.0.1 plug-in
   \item wxWindows Version: wxGTK 2.6.2
\end{itemize}

\paragraph{On Ubuntu 5.10, Breezy Badger}
\begin{itemize}
   \item Compiler: gcc version 4.0.2 20050808 (prerelease) (Ubuntu
4.0.1-4ubuntu9)
   \item IDE Tool: Eclipse 3.1.2, with CDT 3.0.2 plug-in
   \item wxWindows Version: wxGTK 2.6.2
\end{itemize}
