\section{Environment Models}


\subsection{Ephemerides}

\subsubsection{Analytic Ephemeris Model}

\begin{itemize}
     \item For a new body, the user must input the central body by
     choosing from the 9 Planets or the sun.
     %
     \item The user must provide the epoch.
     %
     \item  The user must provide the keplerian elements, in the
     central body centered, MJ2000Eq axis system.
     %
     \item  The user can provide a $\mu$ value for use in the
     solution of the equations of motion.
     %
\end{itemize}

The body should store the users original input for the state and
epoch, and the state and epoch calculated at the last request for
state information. Then, when the next request is made for state
information, the epoch and state from the last request are used as
the input state for next calculation.

\subsection{Atmospheric Density}

\begin{equation}
28K_p + 0.03e^{K_p} = A_p + 100\left(1 - e^{(-0.08A_p)}\right)
\end{equation}

\subsubsection{Jacchia Roberts}

\subsubsection{MSISE-90}

A. E. Hedin, Extension of the MSIS Thermospheric Model into the
Middle and Lower Atmosphere, J. Geophys. Res. 96, 1159, 1991.

Discuss observed vs. adjusted for F10.7 values, also URSI Series D

For testing http://nssdc.gsfc.nasa.gov/space/model/models/msis.html

http://www.agu.org/journals/ja/ja0212/2002JA009430/ \\
go to auxillary material on the left side menu and open the
tables-datasets.doc

Other useful models http://nssdc.gsfc.nasa.gov/space/model/

\subsubsection{Exponential Atmosphere}
